---
title: "p8105_hw2_xz3499"
author: "Xintong Zhao"
date: "2025-10-3"
output: github_document
---

## Set the knitr option
```{r}
knitr::opts_chunk$set(
  echo = TRUE,           
  warning = FALSE,       
  message = FALSE, 
  fig.width = 8,         
  fig.height = 6,        
  fig.align = "center",  
  out.width = "90%",     
  dpi = 300              
)
```

## Load required libraries
```{r}
library(tidyverse)      
library(p8105.datasets) 
library(knitr)          
library(patchwork)      
library(lubridate)      
library(ggthemes)
library(kableExtra)
```

## Set a unified image format and theme
```{r}
theme_set(theme_minimal(base_size = 12) + 
            theme(plot.title = element_text(size = 14, face = "bold"),
                  plot.subtitle = element_text(size = 12),
                  axis.title = element_text(size = 11),
                  axis.text = element_text(size = 10),
                  legend.title = element_text(size = 11),
                  legend.text = element_text(size = 10)))
```

## Problem 1

### 1. Load the dataset
```{r}
data("instacart")
```

### 2. A discription of the dataset
```{r}
glimpse(instacart)
view(instacart)
```

* Size and Structure: 

  * The Instacart dataset contains `r nrow(instacart)` observations of `r ncol(instacart)` variables, representing individual grocery items ordered through the Instacart service. The dataset is structured with each row representing one item within an order, and includes `r length(unique(instacart$order_id))` independent orders, `r length(unique(instacart$user_id))` different users, and `r length(unique(instacart$product_id))` different products. 

* Key variables: 

  * The dataset includes some basic information such as order identification (`order_id`), product information (`product_id`, `product_name`), categorical organization (`aisle_id`, `aisle`, `department_id`, `department`), and user identification (`user_id`). 
  
  * Product classification system: All the products are organized in `r length(unique(instacart$aisle_id))` aisles under `r length(unique(instacart$department_id))` departments.
  
  * Repurchase behavior: Product with `r (mean(instacart$reordered) * 100) %>% round(1)`% are considered repurchase behavior, indicating that the user has a certain degree of brand or product loyalty.
  
  * Shopping cart order: The `add_to_cart_order` records the order in which goods are added to the shopping cart, reflecting the user's shopping decision-making process.
  
  * Time pattern: The dataset contains which day of the week(`order_dow`) and which hour of the day(`order_hour_of_day`) orders are placed, as well as the number of days since the last order was placed(`days_since_prior_order`), which can be used to analyze shopping time references.
  
* Illstrative examples:

  * `r instacart$product_name[1]`(located in `r instacart$aisle[1]` aisle）

  * `r instacart$product_name[50]` (belonging to `r instacart$department[50]` department）

  * `r instacart$product_name[100]` (Added to the shopping cart as the `r instacart$add_to_cart_order[100]` in the order)
  
3. How many aisles are there, and which aisles are the most items ordered from?

The dataset contains `r length(unique(instacart$aisle))` distinct aisles. The top three aisles with the most items ordered are: `r instacart %>% count(aisle, sort = TRUE) %>% slice(1) %>% pull(aisle) `(with `r instacart %>% count(aisle, sort = TRUE) %>% slice(1) %>% pull(n)` items ordered), `r instacart %>% count(aisle, sort = TRUE) %>% slice(2) %>% pull(aisle)` (`r instacart %>% count(aisle, sort = TRUE) %>% slice(2) %>% pull(n)` items), and `r instacart %>% count(aisle, sort = TRUE) %>% slice(3) %>% pull(aisle)` (`r instacart %>% count(aisle, sort = TRUE) %>% slice(3) %>% pull(n)` items).

4. A plot that shows the number of items ordered in each aisle:

```{r}
instacart %>%
  count(aisle) %>%
  filter(n > 10000) %>%
  mutate(aisle = fct_reorder(aisle, n)) %>%
  ggplot(aes(x = n, y = aisle)) +
  geom_col(fill = "steelblue") +
  labs(
    title = "Number of Items Ordered by Aisle (Aisles with >10,000 items)",
    x = "Number of Items Ordered",
    y = "Aisle Name"
  )
```

5. A table showing the three most popular items:

```{r}
target_aisles <- c("baking ingredients", "dog food care", "packaged vegetables fruits")

most_popular_items <- instacart %>%
  filter(aisle %in% target_aisles) %>%
  # Group by aisle and product to count orders per product
  group_by(aisle, product_name) %>%
  summarize(n_ordered = n(), .groups = "drop") %>%
  group_by(aisle) %>%
  slice_max(n_ordered, n = 3) %>%
  arrange(aisle, desc(n_ordered))

most_popular_items %>% 
  kable(caption = "Top Three Most Popular Items in Specified Aisles") 
```

6. A table showing the mean hour of the day
```{r}
mean_hour <- instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = round(mean(order_hour_of_day), 2), .groups = "drop") %>%
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  )

# Add meaningful column names
colnames(mean_hour) <- c("Product Name", "Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")

mean_hour %>% 
  kable(caption = "Average Order Hour (24-hour format) for Pink Lady Apples and Coffee Ice Cream by Day of Week") 
```

## Problem 2

### 1. Load and tidy the datasets
```{r}
zip <- read_csv("zillow_data/Zip Codes.csv") %>% 
  janitor::clean_names()

view(zip)

#Create a borough mapping table
borough_mapping <- data.frame(
  county = c("New York", "Kings", "Queens", "Bronx", "Richmond"),
  borough = c("Manhattan", "Brooklyn", "Queens", "Bronx", "Staten Island")
)

zori <- read_csv("zillow_data/Zip_zori_uc_sfrcondomfr_sm_month_NYC.csv") %>% 
  janitor::clean_names() %>% 
  rename(zip_code = region_name) %>% 
  pivot_longer(
    cols = matches("^x\\d{4}_\\d{2}_\\d{2}$"),
    names_to = "date",
    values_to = "zori"
  ) %>%
  mutate(date = str_remove(date, "^x") %>% 
         str_replace_all("_", "-") %>% 
         ymd())

view(zori)

zillow_data <- zori %>% 
  left_join(
    zip,
    by = "zip_code",
    relationship = "many-to-many"
  ) %>% 
  left_join(borough_mapping, by = "county") %>% 
  select(-county_name)

view(zillow_data)
```

### 2. ZIP code observation frequency
```{r}
# The total number of distinct months
total_months <- zillow_data %>% 
  filter(!is.na(zori)) %>%
  distinct(date) %>% 
  nrow()

# Observation frequency per ZIP cod
zip_obs_freq <- zillow_data %>%
  filter(!is.na(zori)) %>%
  distinct(zip_code, date) %>%
  count(zip_code, name = "n_observations")

# ZIP codes with 116 observations 
obs_full <- zip_obs_freq %>% 
  filter(n_observations == total_months) %>% 
  nrow()

# ZIP codes with fewer than 10 observations
obs_few <- zip_obs_freq %>% 
  filter(n_observations < 10) %>% 
  nrow()

# Calculate percentage coverage
total_zips <- n_distinct(zip_obs_freq$zip_code)
pct_obs_full <- round(obs_full / total_zips * 100, 1)
pct_obs_few <- round(obs_few / total_zips * 100, 1)
```

- The ZIP code observed 116 times : `r obs_full` ( `r pct_obs_full` %)

- ZIP code for observations less than 10 times: `r obs_few` ( `r pct_obs_few` %)

* The reason of some ZIP codes are observed rarely and others observed in each month:

  * Percentile range: Only the listed rent within the 35th to 65th percentile range is calculated. This may affect the observation frequency in some extreme market areas.

  * Market Activity: Postal codes with more active rental markets are more likely to have complete data, while regions with limited rental activities may have sparse observations.

  * Geographical changes: The boundaries of postal codes may have changed over time, or new postal codes may have been created during the research period.

  * Housing Stock characteristics: In regions where self-owned housing is dominant or rental inventory is limited, there will naturally be fewer observations in the data set that focuses on rent.
  
### 3. Average rental price in each borough and year
```{r}
borough_year_avg <- zillow_data %>%
  filter(!is.na(zori), !is.na(borough)) %>%
  mutate(year = year(date)) %>%
  group_by(borough, year) %>%
  summarize(
    avg_zori = round(mean(zori, na.rm = TRUE)),
    .groups = "drop"
  ) %>%
  arrange(borough, year)

borough_year_avg_table <- borough_year_avg %>%
  select(borough, year, avg_zori) %>%
  pivot_wider(
    names_from = year,
    values_from = avg_zori
  )

borough_year_avg_table %>%
  knitr::kable(
    format = "pipe",
    digits = 0
  )
```

- From 2015 to 2024, the rental market in New York City has shown a strong growth trend, with an average increase of approximately `r round(mean(c((4053-3006)/3006, (3126-2493)/2493, (2694-2215)/2215, (2497-1760)/1760), na.rm = TRUE)*100, 0)`%。

- Manhattan increased from $`r 3006` to $`r 4053`, representing a growth of `r round((4053-3006)/3006*100, 1)`%, with a pandemic-related decline to $`r 3091` in 2020 followed by a strong rebound to $`r 3753` in 2022; Brooklyn rose steadily from $`r 2493` to $`r 3126`, growing by `r round((3126-2493)/2493*100, 1)`%; Queens increased from $`r 2215` to $`r 2694`, up `r round((2694-2215)/2215*100, 1)`%, experiencing a significant correction in 2021 with relatively moderate but sustained post-pandemic recovery; The Bronx showed the largest gain, surging from $`r 1760` to $`r 2497`, up `r round((2497-1760)/1760*100, 1)`%, without any decline even in 2020; Staten Island, with records beginning only in 2019, rose from $`r 1978`  to $`r 2536` , growing by `r round((2536-1978)/1978*100, 1)`%.

### 4. NYC rental prices with ZIP codes
```{r}
rent_comparison <- zillow_data %>%
  filter(!is.na(zori), !is.na(borough)) %>%
  # Calculate borough-level monthly averages for cleaner comparison
  group_by(borough, date) %>%
  summarize(
    borough_zori = mean(zori, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  ggplot(aes(x = date, y = borough_zori, color = borough)) +
  geom_line(linewidth = 1.2) +
  geom_point(data = . %>% 
              group_by(borough) %>% 
  # Mark the starting and ending points of time
              filter(date == max(date) | date == min(date) |
  # Mark the pandemic year and the recovery year in particular
                     year(date) %in% c(2020, 2022)), 
            size = 2) +
  # Display the current rental value at the end of time in each boroughs
  geom_text(data = . %>% 
              group_by(borough) %>% 
              filter(date == max(date)), 
            aes(label = paste0("$", round(borough_zori))),
            hjust = -0.2, size = 3, fontface = "bold") +
  labs(
    title = "NYC Rental Prices Comparison (2015-2024)",
    x = "Year",
    y = "Average Rental Prices ($)",
    color = "Borough"
  ) +
  scale_y_continuous(
    labels = scales::dollar_format(),
    limits = c(1000, 4500) 
  ) +
  scale_x_date(
    date_breaks = "1 year", 
    date_labels = "%Y",
    expand = expansion(mult = c(0.05, 0.1)) 
  ) +
  scale_color_brewer(palette = "Set1") +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(face = "bold"),
    panel.grid.major = element_line(color = "grey90"),
    panel.grid.minor = element_line(color = "grey95")
  )

rent_comparison
```

As shown in the figure, all boroughs have shown an overall upward trend. Obvious price fluctuations can be seen around 2020, and prices have gradually rebounded in the post-pandemic period. The relative positions among the various boroughs remained stable throughout the period. Manhattan is at the top of the picture, indicating its highest rental level, while Staten Island is at the bottom, with a significant gap between the two.

### 5. Average rental price within each ZIP code over each month in 2023
```{r}
zip_level_data <- zillow_data %>%
  filter(year(date) == 2023, !is.na(zori), !is.na(borough)) %>%
  group_by(zip_code, borough) %>%
  summarize(
    avg_rent_2023 = mean(zori, na.rm = TRUE),
    n_months = n(),
    .groups = "drop"
  ) %>%
  filter(n_months >= 6)

borough_stats <- zip_level_data %>%
  group_by(borough) %>%
  summarize(
    median_rent = median(avg_rent_2023),
    mean_rent = mean(avg_rent_2023),
    q1 = quantile(avg_rent_2023, 0.25),
    q3 = quantile(avg_rent_2023, 0.75),
    n_zipcodes = n(),
    .groups = "drop"
  )

dist_comparison_2023 <- zip_level_data %>%  
  ggplot(aes(x = fct_reorder(borough, avg_rent_2023, .fun = median), 
             y = avg_rent_2023, 
             fill = borough)) +
  geom_boxplot(alpha = 0.7, outlier.shape = NA, width = 0.6) +
  geom_jitter(width = 0.2, alpha = 0.5, size = 1.2, aes(color = borough)) +
  geom_text(
    data = borough_stats,  
    aes(x = borough, y = median_rent, 
        label = paste0("Median\n$", round(median_rent))),
    vjust = -0.8, size = 3, fontface = "bold"
  ) +
  labs(
    title = "Distribution of Average Rental Prices (2023)",
    x = "Borough",
    y = "Average Rental Prices($)",
    fill = "Borough",
    color = "Borough"
  ) +
  scale_y_continuous(labels = scales::dollar_format()) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, face = "bold"),
    legend.position = "none",
    plot.title = element_text(face = "bold")
  )

dist_comparison_2023
```

The rents in various boroughs of New York show a clear hierarchical structure: Manhattan is at the top of the map, with the widest distribution and a significant right bias. Brooklyn and Queens form the mid-range market, with Queens having a relatively concentrated distribution. Staten Island and the Bronx form a relatively affordable housing market, with Staten Island having a more compact distribution.

### 6. Combine the two plots
```{r}
combined_plot <- rent_comparison + dist_comparison_2023 +
  plot_annotation(
    title = "Comprehensive Analysis of NYC Rental Market Dynamics",
    tag_levels = 'A',
    theme = theme(
      plot.title = element_text(face = "bold", size = 14, hjust = 0.5)
    )
  ) +
  plot_layout(widths = c(1.5, 1)) 

combined_plot

ggsave(
  "results/nyc_rental_prices_comprehensive_analysis.png",
  combined_plot,
  width = 16,
  height = 8, 
  dpi = 300,
  bg = "white"
)
```

